Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2013-01-14 chi2013-webbox.tex
foo
latest

\begin{abstract}
The information processing capabilities of humans enable them to opportunistically draw and integrate knowledge from nearly any information source.  However, the integration of digital, structured data from diverse sources remains difficult, due to problems of heterogeneity that arise when data modelled separately are brought together.  In this paper, we present an investigation of the feasibility of extending Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, with the objective of allowing  users to take advantage of the emerging ecosystems of structured data currently becoming available.  In this study, we conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources. Observations from these pre-studies led to \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.  Our lab study showed, first, that participants readily understood the new interaction mechanisms which were introduced.  Secondly, participants made more carefully justified decisions, even while weighing a greater number of factors, moreover expending less effort, during subjective-choice tasks when using DataPalette, than with a control set-up.

@@

\subsubsection{Results}
While most of the data records examined in PS.2 were small and relatively simple, nearly one in each category was substantially larger and more complex. In particular, Amazon, Soundcloud, Twitter and Weather Underground were the most complex in each of their respective categories. Mapping equivalent properties among the instance records for each category revealed little overall overlap, except among the smallest records (which had few properties to begin with).  This finding suggests that consolidating such records would substantially increase the amount of information held by each.  Some of the unique properties were primarily for internal use (such as service-specific IDs).  Some, however, contained useful properties which were simply absent in alternate sources.  For example, among the 4 music data providers examined, only Spotify listed the source album that featured a particular song.  Such omissions revealed that data APIs often were reflective of each provider's own specific needs, necessitating the integration of information across multiple sources to create more complete and useful representations.

@@

When asked to rate the DP interface in the exit survey, all participants responded that they felt they understood the system, with sixteen out of the twenty participants (80\%) responding that they agreed \emph{very strongly} that they felt they understood it.  Eighteen (90\%) reported that they thought  DP was \emph{easy}, as well as \emph{useful}, with one person being neutral on each.  As to the statement ``DP is confusing'',  11 (55\%) strongly disagreed, 6 (30\%) disagreed, and 3 (15\%) agreed.  On reviewing the time trials, we discovered that the 3 participants who rated DP as confusing completed the trials significantly more slowly than the rest of the group. The survey results were visualised in Figure \ref{fig:survey}.

@@

%% -> discussion and limitations
\subsection{Design Limitations}
In addition to the aforementioned feature requests, we observed a number of potential areas for improvement in the design of the DP interface.  One such area pertained to window layout and screen real-estate management: participants were prone to opening up a large number of windows, including maps and charts.  When their workspaces started to fill up, we found that participants spent a lot of time arranging windows, and that they tended to forget what each window represented.  We plan to address these issues in the next version of DataPalette.

A key design decision we made at the outset was to target only the most common forms of heterogeneity observed in PS.2.  To expand DP's integration capabilities, we are considering a number of new interaction affordances for facilitating field combination and splitting, and unit of measure reconciliation.  This would cover a majority of the structural and modelling-oriented heterogeneity cases observed in DP.2.  Extending the ``Do What I Mean'' capabilities of the visualisation features to be unit-of-measure-aware, we believe, would be valuable to users.  A core problem with realising such an capability, however, is that few sources explicitly annotated data values with measurement units.

@@

>>>>

\begin{abstract}
The information processing capabilities of humans enable them to opportunistically draw and integrate knowledge from nearly any information source.  However, the integration of digital, structured data from diverse sources remains difficult, due to problems of heterogeneity that arise when data modelled separately are brought together.  In this paper, we present an investigation of the feasibility of extending Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, with the objective of allowing  users to take advantage of the emerging ecosystems of structured data currently becoming available.  In this study, we conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources. Observations from these pre-studies led to \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.  Our lab study showed that participants readily understood the new interaction mechanisms which were introduced.  Participants made more carefully justified decisions, even while weighing a greater number of factors, moreover expending less effort, during subjective-choice tasks when using DataPalette, than with a control set-up.

@@

\subsubsection{Results}
While most of the data records examined in PS.2 were small and relatively simple, nearly one in each category was substantially larger and more complex. In particular, Amazon, Soundcloud, Twitter and Weather Underground were the most complex in each of their respective categories. Mapping equivalent properties among the instance records for each category revealed little overall overlap, except among the smallest records (which had few properties to begin with).  This finding suggested that consolidating such records would substantially increase the amount of information held by each.  Some of the unique properties were primarily for internal use (such as service-specific IDs).  Some, however, contained useful properties which were simply absent in alternate sources.  For example, among the 4 music data providers examined, only Spotify listed the source album that featured a particular song.  Such omissions revealed that data APIs often were reflective of each provider's own specific needs, necessitating the integration of information across multiple sources to create more complete and useful representations.

@@

When asked to rate the DP interface in the exit survey, all participants responded that they felt they understood the system, with sixteen out of the twenty participants (80\%) responding that they agreed \emph{very strongly}.  Eighteen (90\%) reported that they thought  DP was \emph{easy}, as well as \emph{useful}, with one person being neutral on each.  As to the statement ``DP is confusing'',  11 (55\%) strongly disagreed, 6 (30\%) disagreed, and 3 (15\%) agreed.  On reviewing the time trials, we discovered that the 3 participants who rated DP as confusing completed the trials significantly more slowly than the rest of the group. (See  Figure \ref{fig:survey}.)

@@

%% -> discussion and limitations
\subsection{Design Limitations}
In addition to the aforementioned feature requests, we observed a number of potential areas for improvement in the design of the DP interface.  One such area pertained to window layout and screen real-estate management: participants were prone to opening up a large number of windows, including maps and charts.  When their workspaces started to fill up, we found that participants spent a lot of time arranging windows, and that they tended to forget what each window represented.  We plan to address these issues in the next version of DP.

A key design decision we made at the outset was to target only the most common forms of heterogeneity observed in PS.2.  To expand DP's integration capabilities, we are considering a number of new interaction affordances for facilitating field combination and splitting, and unit of measure reconciliation.  This would cover a majority of the structural and modelling-oriented heterogeneity cases observed in PS.2.  Extending the ``Do What I Mean'' capabilities of the visualisation features to be unit-of-measure-aware, we believe, would be valuable to users.  A core problem with realising such an capability, however, is that few sources explicitly annotated data values with measurement units.

@@

\newpage
